### 2012-10-27 paper_draft.tex
PNG-versions of the graphs, updated most stats (except # UserSNPs, still counting), moved tables, inserted new graphs

\begin{table}
\begin{tabular}{|l|l|l|}
\hline
& Turkey's HSD & \\ \cline{2-3}
& Mean difference & SE \\ \hline
\textbf{Motivation for sharing genotypings in participants} & & \\
\textbf{who are already genotyped} & & \\ \hline 
... curious & 1.159 & 0.193 \\ 
... want to help scientists & 0.465 & 0.128 \\
... for personal benefits & 0.448 & 0.183 \\ \hline
\textbf{Motivation for not sharing in participants} & & \\ 
\textbf{who are not planning to get genotyped} & & \\ \hline
... fear of discrimination & 1.06 & 0.195 \\
... breach of privacy & 0.821 & 0.225 \\
... fear of personalized advertising & 0.848 & 0.208 \\ 
... negative consequences for family members & 0.733 & 0.21 \\ \hline 
\end{tabular}
\caption{Differences in terms of motivation to share genotypings with the public in survey-participants who already received a genotyping compared to participants who are not planning to getting genotyped. }
\label{tab:motivations1}
\end{table}

\begin{table}
\begin{tabular}{|l|l|l|}
\hline
& Turkey's HSD & \\ \cline{2-3}
& Mean difference & SE \\ \hline
\textbf{Motivation for sharing genotypings in participants} & & \\
\textbf{who would share with their DTC provider} & & \\ \hline
... curiosity & 1.99 & 0.321 \\
... want to help science & 1.57 & 0.199 \\
... for personal benefits & 0.951 & 0.308 \\ \hline
\textbf{Motivation for sharing genotypings in participants} & & \\
\textbf{who would not share with their DTC provider} & & \\ \hline
... fear of discrimination & 1.52 & 0.322 \\
... fear of consequences for family members & 1.146 & 0.32 \\
... fear of personalized advertising & 1.112 & 0.357 \\ \hline
\end{tabular}
\caption{Differences in terms of motivations to share genotyping-data, comparison between participants who would share their genotyping data with their DTC provider with participants who would not share their data with their DTC provider.}
\label{tab:motivations2}
\end{table}

@@

\subsection*{Sharing genotypic information}
Currently users can upload their genotyping results from the companies \textit{23andMe}, \textit{deCODEme }and \textit{FamilyTreeDNA} via a webinterface to the openSNP 
project. There is experimental support for uploading exomes in the VCF format \cite{Danecek01082011}, as \textit{23andMe} recently started exome sequencing for its customers. So far only the SNPs of the exome data sets are visualized on openSNP, but the downloads include all variation found in the exome.
The uploaded data is published under the Creative Commons Zero-license, 
which - in accordance with the Panton Principles \cite{10.1371/journal.pbio.1001195} - 
allows a complete reuse of the data without any constraints.
Between the start of openSNP on 09/27/2011 and 12/18/2012, 214 people have signed 
up with openSNP, and 79 genotyping files were made available. The openSNP 
database lists 69,486,471 genotypes which are distributed over 1,938,603 unique SNPs.
Figure \ref{Figure1_label} depicts the increase of users and genotyping files over time.\bastian{update all numbers}

@@

In the same timeframe as above, all users combined have 
entered a total of 675 variations on 47 different phenotypes with those variations being 
the different values on a given trait or phenotype. See figure \ref{Figure1_label} for the increase of phenotypic information over time.

The mean number of users that have entered their variations for a single phenotype 
is 14.36 (SD 12.65), the median is 10. The distribution of how many users have 
entered their data per phenotype can be seen in figure \ref{Figure2_label}. The phenotype provided 
by the most users is the eye color, which has been entered by 54 users. There are 
two phenotypes which have so far only been provided by a single user: 
the SAT Writing score and triglyceride-levels.\bastian{update all numbers and graphs}

\subsection*{Connection to external services}
In order to provide users with relevant information on their respective genotypes, openSNP scans databases of the scientific literature for specific SNPs. 
A total number of 15,229 documents \bastian{number needs to get updated}relevant to the SNPs listed in openSNP could be found in the publication databases of Mendeley, the Public Library of Science and in the crowdsourced SNPedia. 
Of the primary literature, 25 \% are released in open access journals and can be accessed free of charge (Figure \ref{Figure3_label}). For usability reasons, 
SNPs are ranked by the amount of information gathered through the external services. The external services themselves are ranked by how easily non-scientists can understand information 
from these sources and available this information. The SNPedia entries are given the highest impact, as those are already manually curated and summarized in plain English, followed by open access publications out of 
the Public Library of Science. Lowest values are given to the Mendeley results, as the publications listed there are for the most part not freely available without subscriptions or one-time payments. 
An entry on SNPedia is valued 2.5 times as high as a PLoS publication and 5 times as high as a Mendeley entry.

@@

\section*{Figure Legends}
\begin{figure}[!ht]
	\begin{center}
		\includegraphics[scale=0.35]{chart_growth.png}
	\end{center}
	\caption{
	{\bf Growth of openSNP.} The increase in numbers for users, genotyping-files, phenotypes and their variation from 27.09.2011 to 16.12.2011 is shown.} 
	\label{Figure1_label}
\end{figure}

\begin{figure}[!ht]
	\begin{center}
		\includegraphics[scale=0.40]{histogram_phenotypes.png}
	\end{center}
	\caption{
	{\bf Histogram of users/phenotype-distribution.} The x-axis shows the minimum number of users who provide information for a phenotype, the y-axis shows how many phenotypes have at least that many users.}
	\label{Figure2_label}
\end{figure}

\begin{figure}[!ht]
	\begin{center}
		\includegraphics[scale=0.50]{paper_distribution.png}
	\end{center}
	\caption{
	{\bf Distribution of external information gathered for SNPs in the openSNP-database.} Data on PLoS and SNPedia is openly available for every user. Publications on Mendeley are either Open Access (OA) or Closed Access (CA).} 
	\label{Figure3_label}
\end{figure}

\begin{figure}[!ht]
	\begin{center}
		\includegraphics[scale=0.60]{uml_diagram.png}
	\end{center}
	\caption{
	{\bf Flow of data inside openSNP.} External databases and user-provided data are used as input. Output of data is done using the website, the \emph{Distributed Annotation System} and a JSON-API.} 
	\label{Figure4_label}
\end{figure}

@@

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

>>>>

@@

\subsection*{Sharing genotypic information}
Currently users can upload their genotyping results from the companies \textit{23andMe}, \textit{deCODEme }and \textit{FamilyTreeDNA} via a webinterface to the openSNP 
project. There is experimental support for uploading exomes in the VCF format \cite{Danecek01082011}, as \textit{23andMe} recently started exome sequencing for its customers. So far only the SNPs of the exome data sets are visualized on openSNP, but the downloads include all variation found in the exome.
The uploaded data is published under the Creative Commons Zero-license, 
which - in accordance with the Panton Principles \cite{10.1371/journal.pbio.1001195} - 
allows a complete reuse of the data without any constraints.
Between the start of openSNP on 09/27/2011 and 10/27/2012, 633 people have signed 
up with openSNP, and 270 genetic datasets were made available. The openSNP 
database lists X genotypes which are distributed over 2,140,643 unique SNPs.
Figures \ref{Figure1_label} and \ref{Figure2_label} depicts the increase of users and genotyping files over time.\bastian{update all numbers}

@@

In the same timeframe as above, all users combined have 
entered a total of 4743 variations on 130 different phenotypes with those variations being 
the different values on a given trait or phenotype. The mean number of users that have entered their variations for a single phenotype 
is 36.48. The distribution of how many users have 
entered their data per phenotype , compared to the amount of unique phenotypes, can be seen in figure \ref{pheno}. The phenotype provided by the most users is "eye color", for which 207 users entered their phenotype.

\subsection*{Connection to external services}
In order to provide users with relevant information on their respective genotypes, openSNP scans databases of the scientific literature for specific SNPs. 
A total number of 15,229 documents \bastian{number needs to get updated}relevant to the SNPs listed in openSNP could be found in the publication databases of Mendeley, the Public Library of Science and in the crowdsourced SNPedia. 
Of the primary literature, 25 \% are released in open access journals and can be accessed free of charge. For usability reasons, 
SNPs are ranked by the amount of information gathered through the external services. The external services themselves are ranked by how easily non-scientists can understand information 
from these sources and available this information. The SNPedia entries are given the highest impact, as those are already manually curated and summarized in plain English, followed by open access publications out of 
the Public Library of Science. Lowest values are given to the Mendeley results, as the publications listed there are for the most part not freely available without subscriptions or one-time payments. 
An entry on SNPedia is valued 2.5 times as high as a PLoS publication and 5 times as high as a Mendeley entry.

Users are also able to link their Fitbit-accounts to their user-accounts. Fitbit is a commercial service which lets their customers track their BMI, movement-data and sleep data. This data can be linked to openSNP to give interested researchers an automatically maintained dataset of body- and sleep-developments over time.

@@

\section*{Figure Legends}
\begin{figure}[!ht]
	\begin{center}
		\includegraphics[scale=0.35]{25_10_2012_Graphs/users.png}
	\end{center}
	\caption{
	{\bf Growth of openSNP-user-accounts.} The increase in numbers for users from 27.09.2011 to 27.10.2012 is shown.} 
	\label{Figure1_label}
\end{figure}
\begin{figure}[!ht]
	\begin{center}
		\includegraphics[scale=0.35]{25_10_2012_Graphs/genotypes.png}
	\end{center}
	\caption{
	{\bf Growth of available genotypings.} The increase in numbers for genotyping-files from 27.09.2011 to 27.10.2012 is shown.} 
	\label{Figure2_label}
\end{figure}
\begin{figure}[!ht]
	\begin{center}
		\includegraphics[scale=0.40]{25_10_2012_Graphs/phenotypes_vs_userphenotypes.png}
	\end{center}
	\caption{
	{\bf Development of unique phenotypes and phenotypic information over time.} The x-axis shows the time-frame from start of the project until October 2012, the left y-axis shows how many unique phenotypes have been entered, and the right y-axis shows the amount of phenotypes users entered.}
	\label{pheno}
\end{figure}

\begin{figure}[!ht]
	\begin{center}

	\end{center}
	\caption{
	{\bf Flow of data inside openSNP.} External databases and user-provided data are used as input. Output of data is done using the website, the \emph{Distributed Annotation System} and a JSON-API.} 
	\label{Figure4_label}
\end{figure}

@@

\section*{Tables}
\begin{table}
\begin{tabular}{|l|l|l|}
\hline
& Turkey's HSD & \\ \cline{2-3}
& Mean difference & SE \\ \hline
\textbf{Motivation for sharing genotypings in participants} & & \\
\textbf{who are already genotyped} & & \\ \hline 
... curious & 1.159 & 0.193 \\ 
... want to help scientists & 0.465 & 0.128 \\
... for personal benefits & 0.448 & 0.183 \\ \hline
\textbf{Motivation for not sharing in participants} & & \\ 
\textbf{who are not planning to get genotyped} & & \\ \hline
... fear of discrimination & 1.06 & 0.195 \\
... breach of privacy & 0.821 & 0.225 \\
... fear of personalized advertising & 0.848 & 0.208 \\ 
... negative consequences for family members & 0.733 & 0.21 \\ \hline 
\end{tabular}
\caption{Differences in terms of motivation to share genotypings with the public in survey-participants who already received a genotyping compared to participants who are not planning to getting genotyped. }
\label{tab:motivations1}
\end{table}

\begin{table}
\begin{tabular}{|l|l|l|}
\hline
& Turkey's HSD & \\ \cline{2-3}
& Mean difference & SE \\ \hline
\textbf{Motivation for sharing genotypings in participants} & & \\
\textbf{who would share with their DTC provider} & & \\ \hline
... curiosity & 1.99 & 0.321 \\
... want to help science & 1.57 & 0.199 \\
... for personal benefits & 0.951 & 0.308 \\ \hline
\textbf{Motivation for sharing genotypings in participants} & & \\
\textbf{who would not share with their DTC provider} & & \\ \hline
... fear of discrimination & 1.52 & 0.322 \\
... fear of consequences for family members & 1.146 & 0.32 \\
... fear of personalized advertising & 1.112 & 0.357 \\ \hline
\end{tabular}
\caption{Differences in terms of motivations to share genotyping-data, comparison between participants who would share their genotyping data with their DTC provider with participants who would not share their data with their DTC provider.}
\label{tab:motivations2}
\end{table}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}
